---
title: "Lauren Neudorf Atlanta Hawks Data Analyst Interview"
author: "Lauren Neudorf"
date: "2023-06-02"
output:
  word_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r read}
setwd("~/Desktop/etc./notes/Stats Method 1/HW")
library(readxl)
nbaplayers=read_excel("nbaplayers.xlsx")

```
Part 1. A (a):

#Filtering/minimizing df for duke players, then for those Duke players drafted before or during 2000, then counting the frequency of the teams. 
```{r pt1A a}
duke=nbaplayers[nbaplayers$college=="Duke",]
duke$year=as.numeric(duke$year)
duke=duke[,1:7]
duke=na.omit(duke)
yearorder=duke[order(duke$year),]

b42000=yearorder[1:17,]

library(vctrs)
vec_count(b42000$team)

```
Part 1 A (b):

First, I created a subset of the original df with just even draft years, then I created a subset of that with player names that have a "D", and sorted alphabetically. I then created a third sub dataframe with just first names that start with "D", this df has 110 rows, i.e. there are 110 players that fit this criteria. 

```{r pt1A b}

copydf=nbaplayers
copydf$year=as.numeric(copydf$year)
newdata=subset(copydf, copydf$year=="1990"|copydf$year=="1992"|copydf$year=="1994"|copydf$year=="1996"|copydf$year=="1998"|copydf$year=="2000"|copydf$year=="2002"|copydf$year=="2004"|copydf$year=="2006"|copydf$year=="2008"|copydf$year=="2010"|copydf$year=="2012"|copydf$year=="2014"|copydf$year=="2016"|copydf$year=="2018"|copydf$year=="2020")
library(data.table)
dnames=newdata[newdata$player %like% "D",]
dnames=dnames[order(dnames$player),]
dnames=dnames[12:121,]

```

Part 1 B:


First I sorted the df by team and within team, by draft year. Then I created a subset of the df for just first round picks #1-30. Then I created an example of 
```{r pt1B}

nbaplayers=nbaplayers[order(nbaplayers$team,nbaplayers$year),]
first.round=subset(nbaplayers,overall_pick<=30)

atl=subset(first.round, team=="ATL")
plot(atl$year,atl$overall_pick)
library(ggplot2)
ggplot(first.round, aes(x =year, y = overall_pick, color = team)) +
  geom_point() +
  facet_wrap( ~ team)



```
```{r pt2A}

avgWS=read_excel("Avg WS.xlsx")

#preliminary plot
plot(avgWS$overall_pick,avgWS$`Average of win_shares`)


avgWS$overall_pick=as.numeric(avgWS$overall_pick)
avgWS$`Average of win_shares`=as.numeric(avgWS$`Average of win_shares`)
y=avgWS$`Average of win_shares`
x=avgWS$overall_pick
library(caret)

#apply a local regression, get predicted values i.e the "smoothed out" curve
modavgWS=loess(y ~ x, data=avgWS, span=.75)
assigned.values=predict(modavgWS) 
assigned.valu


#scatterplot with regression line overlaid
plot(avgWS$overall_pick, avgWS$`Average of win_shares`, pch=19, main='Loess Regression Model', xlim=c(0,61), ylim=c(-5,100))

lines(assigned.values, x=avgWS$overall_pick, col='purple')

#create df of these assigned values per pick for future use
Pick.number=as.data.frame(avgWS$overall_pick)
assignedvals=as.data.frame(assigned.values)
values.per.pick=cbind(Pick.number,assignedvals)
values.per.pick


```
```{r pt2b}

plot(values.per.pick$assigned.values, avgWS$`Average of win_shares`)



```

##Code Resources Used links

Used to confirm how to use local regression in R:

https://www.statology.org/loess-regression-in-r/

Used to find alternative method to original selected method: 

https://cran.r-project.org/web/packages/lessR/vignettes/Extract.html

To plot the grouped data:

https://stackoverflow.com/questions/15546468/plot-grouped-data-in-r

